Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-26 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
trying to trim text

The paper is organized as follows. In Sec.~\ref{sec:review_theories}, we briefly describe the four
higher-curvature modified gravity theories for which we peform the ringdown test.
In Sec.~\ref{sec:method} we build our parameterized IMR model for nonprecessing-spin compact-object
binaries making use of the ParSpec framework. After reviewing the Bayesian inference method,
in Sec.~\ref{sec:pe}, we motivate our selection of GW events from the LVK catalog, and also
discuss the range of validity of our analyses. In particular, we discuss the impact on
our results of the assumptions underlying the ParSpec framework, and the fact that
our modified gravity theories have to be interpreted as EFTs. In Sec.~\ref{sec:results},
we present our results obtained by applying Bayesian analysis on the LVK data of
GW150914 and GW1200129, and discuss how we set the upper limits
on the fundamental length-scales $\ell_{\rm th}$. Finally, in Sec.~\ref{sec:conclusions} we
summarize our findings, and discuss how to make our framework more robust, in view
of stronger GW events in upcoming GW observational runs, by including  physical effects currently absent in our
study (e.g., precessing-spins and eccentricity). In the Appendix~\ref{app:map_details}
we provide details in calculating the non-GR QNM frequencies, when using ParSpec,
for the four modified gravity theories under consideration.
%
Henceforth, unless otherwise specified, we work in geometric units $G = 1 = c$.

@@

In the end, the $\pSEOB$ waveform model with theory-specific
QNMs has only $\ell_{\rm th}$ as a free beyond-GR parameter.
%
We emphasize that our procedure is different from that of
Ref.~\cite{Carullo:2021dui} which, for a given value of $p$, varied all $\ell_{\rm th}$,
$\delta\omega^{(j)}$, and $\delta\tau^{(j)}$ parameters, and then proceeded
to use the posteriors on $\ell_{\rm th}$ to draw conclusions to any theory that predict
corrections at a given $p$.
%
This is a important difference between our work and Ref.~\cite{Carullo:2021dui}.
In particular, as we will see in Sec.~\ref{sec:results}, adding theory-specific
information to the ParSpec coefficients \emph{can lead to different interpretations
of the final bounds on $\ell_{\rm th}$, even for different theories that predict the same
value of the exponent $p$.}

@@

In particular, we found, that within the interpretation of these theories as EFTs and the
region of validity of the ParSpec framework, the fundamental length-scale of dCS gravity is bound as
$\ell_{\rm dCS} \leqslant 34.5$~km, at 90\% credible level, when stacking the posteriors
of GW150914 and GW200129. This is the first GW-alone constrain on this theory.
%
In contrast, we could not place any bounds on the \hs{fundamental} length-scale
% \abcomm{Maybe
% this is common in the literature, but I would call it length-scale instead of coupling
% constant. What do you think?}~\hscomm{I agree. I introduced ``length-scale'' at a few other place in the text.}
of EdGB gravity $\ell_{\rm EdGB}$.
%
This dichotomy between the two theories has a counterpart with works that considered
the inspiral part of the GW signal alone~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}.
%
Using data of the LVK BBHs, it was found that the posterior distributions for
deviations from GR were uninformative in dCS gravity, but not in EdGB gravity.
We emphasize that both those theories (and the cubic EFT of GR also studied here) all predict the same exponent $p$ in ParSpec.
%
Hence, our results show how the inclusion of theory-specific information into the ParSpec framework
can result in different outcomes for different theories, even if they predict the same value of $p$.

>>>>

The paper is organized as follows. In Sec.~\ref{sec:review_theories}, we briefly describe the four
higher-curvature modified gravity theories for which we peform the ringdown test.
In Sec.~\ref{sec:method} we build our parameterized IMR model for nonprecessing-spin compact-object
binaries making use of the ParSpec framework. After reviewing the Bayesian inference method,
in Sec.~\ref{sec:pe}, we motivate our selection of GW events from the LVK catalog, and also
discuss the range of validity of our analyses. In particular, we discuss the impact on
our results of the assumptions underlying the ParSpec framework, and the fact that
our modified gravity theories have to be interpreted as EFTs. In Sec.~\ref{sec:results},
we present our results obtained by applying Bayesian analysis on the LVK data of
GW150914 and GW1200129, and discuss how we set the upper limits
on the fundamental length-scales $\ell_{\rm th}$. Finally, in Sec.~\ref{sec:conclusions} we
summarize our findings, and discuss how to make our framework more robust, in view
of stronger GW events in upcoming GW observational runs, by including  physical effects currently absent in our
study (e.g., precessing-spins and eccentricity). In the Appendix~\ref{app:map_details}
we provide details in calculating the non-GR QNM frequencies, when using ParSpec,
for the modified gravity theories under consideration.
%
Henceforth, unless otherwise specified, we work in geometric units $G = 1 = c$.

@@

In the end, the $\pSEOB$ waveform model with theory-specific
QNMs has only $\ell_{\rm th}$ as a free beyond-GR parameter.
%
We emphasize that our procedure is different from that of
Ref.~\cite{Carullo:2021dui} which, for a given value of $p$, varied all $\ell_{\rm th}$,
$\delta\omega^{(j)}$, and $\delta\tau^{(j)}$ parameters, and then proceeded
to use the posteriors on $\ell_{\rm th}$ to draw conclusions to any theory that predict
corrections at a given $p$.
%
This is a important difference between our work and Ref.~\cite{Carullo:2021dui}.
In particular, as we will see in Sec.~\ref{sec:results}, adding theory-specific
information to the ParSpec coefficients \emph{can lead to different interpretations
of the bounds on $\ell_{\rm th}$, even for different theories that predict the same
value of the exponent $p$.}

@@

In particular, we found, that within the interpretation of these theories as EFTs and the
region of validity of the ParSpec framework, the fundamental length-scale of dCS gravity is bound as
$\ell_{\rm dCS} \leqslant 34.5$~km, at 90\% credible level, when stacking the posteriors
of GW150914 and GW200129. This is the first GW-alone constrain on this theory.
%
In contrast, we could not place any bounds on the fundamental length-scale
% \abcomm{Maybe
% this is common in the literature, but I would call it length-scale instead of coupling
% constant. What do you think?}~\hscomm{I agree. I introduced ``length-scale'' at a few other place in the text.}
of EdGB gravity $\ell_{\rm EdGB}$.
%
This dichotomy between the two theories has a counterpart with works that considered
the inspiral part of the GW signal alone~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}.
%
Using data of the LVK BBHs, it was found that the posterior distributions for
deviations from GR were uninformative in dCS gravity, but not in EdGB gravity.
We emphasize that both those theories (and the cubic EFT of GR also studied here) all predict the same exponent $p$ in ParSpec.
%
Hence, our results show how the inclusion of theory-specific information into the ParSpec framework
can result in different outcomes for different theories, even if they predict the same value of $p$.
